---
title: "Mean difference estimation"
author: "Thejasvi"
date: "26/2/2021"
output: html_document
---

This notebook will run bayesian posterior distribution analysis of the linear mixed models that have been inspected in the previous notebook titled 'horseshoe-bat-analysis.Rmd'.
  
```{r loading, include=FALSE, echo=TRUE}
library(arm)
library(coda)
library(lme4)

d <- read.csv('timeordered_indcall.csv')

```

## tFM duration 

```{r tfmdurn}
tfmdurn.nona <- d[!is.na(d$tfm_duration),]
tfmdurn.lmm <- lmer(tfm_duration~multi_bat+(1|rechour), data=tfmdurn.nona)

# - now run bayesian posterior distbn estimatin 
tfmdurn.sim <- sim(tfmdurn.lmm, n.sim=5000)
pctile.range <- 0.06
tfmdurn.meandiff.range <- quantile(tfmdurn.sim@fixef[,2], c(pctile.range*0.5,1-(pctile.range*0.5)))
tfmdurn.meandiff.hpd <- HPDinterval(as.mcmc(tfmdurn.sim@fixef[,2]), prob=0.93)
tfmdurn.singlebt.hpd <- HPDinterval(as.mcmc(tfmdurn.sim@fixef[,1]), prob=0.93)

# - quantify the relative increase in tfm duration between single and multiple bats:
tfmdurn.relative.increase <- (tfmdurn.sim@fixef[,1]+tfmdurn.sim@fixef[,2])/tfmdurn.sim@fixef[,1]
tfmdurn.relincrease.hpd <- round(HPDinterval(as.mcmc(tfmdurn.relative.increase), prob=0.93),2)

hist(tfmdurn.relative.increase)
```
There is a relative increase in tFM duration is between `r tfmdurn.relincrease.hpd[1]`-`r tfmdurn.relincrease.hpd[2]` times that of single bat calls. 

## iFM duration 
```{r ifmdurn}
ifmdurn.nona <- d[!is.na(d$ifm_duration),]
ifmdurn.lmm <- lmer(ifm_duration~multi_bat+(1|rechour), data=ifmdurn.nona)

# - now run bayesian posterior distbn estimatin 
ifmdurn.sim <- sim(ifmdurn.lmm, n.sim=5000)
ifmdurn.meandiff.range <- quantile(ifmdurn.sim@fixef[,2], c(pctile.range*0.5,1-(pctile.range*0.5)))
ifmdurn.meandiff.hpd <- HPDinterval(as.mcmc(ifmdurn.sim@fixef[,2]), prob=0.93)
ifmdurn.singlebt.hpd <- HPDinterval(as.mcmc(ifmdurn.sim@fixef[,1]), prob=0.93)

# - quantify the relative increase in tfm duration between single and multiple bats:
ifmdurn.relative.increase <- (ifmdurn.sim@fixef[,1]+ifmdurn.sim@fixef[,2])/ifmdurn.sim@fixef[,1]
ifmdurn.relincrease.hpd <- round(HPDinterval(as.mcmc(ifmdurn.relative.increase), prob=0.93),2)

hist(ifmdurn.relative.increase)
```
There is a relative change in multi-bat iFM duration between `r ifmdurn.relincrease.hpd[1]`-`r ifmdurn.relincrease.hpd[2]` times that of single-bat calls. 

## CF duration 
```{r cfdurn}
cfdurn.nona <- d[!is.na(d$cf_duration),]
cfdurn.nona$logcfdurn <- log(cfdurn.nona$cf_duration)
cfdurn2.lmm <- lmer(logcfdurn~multi_bat+(1|rechour), data=cfdurn.nona)



# - now run bayesian posterior distbn estimatin 
cfdurn.sim <- sim(cfdurn2.lmm, n.sim=5000)
cfdurn.meandiff.range <- quantile(cfdurn.sim@fixef[,2], c(pctile.range*0.5,1-(pctile.range*0.5)))
cfdurn.meandiff.hpd <- HPDinterval(as.mcmc(cfdurn.sim@fixef[,2]), prob=0.93)
cfdurn.singlebt.hpd <- HPDinterval(as.mcmc(cfdurn.sim@fixef[,1]), prob=0.93)

# - quantify the relative increase in tfm duration between single and multiple bats:
cfdurn.relative.increase <- (exp(cfdurn.sim@fixef[,1]+cfdurn.sim@fixef[,2]))/exp(cfdurn.sim@fixef[,1])
cfdurn.relincrease.hpd <- round(HPDinterval(as.mcmc(cfdurn.relative.increase), prob=0.93),2)

hist(cfdurn.relative.increase, xlab='Relative increase to single bat context', main='Multi-bat increase')
abline(v=cfdurn.relincrease.hpd)
```
There is a relative change in multi-bat CF duration between `r cfdurn.relincrease.hpd[1]`-`r cfdurn.relincrease.hpd[2]` times that of single-bat calls. 




