---
title: "Horseshoe bat data analysis"
author: "Thejasvi"
date: "2/25/2021"
output: html_document
---

## Model generation and diagnostics for individul call analysis 
Here we will be looking at the individual call analysis and estimating the mean difference of various parameters between single and multiple bat calls. 

* TODO: 
1. To check the residuals from the random effects -- see how to do the complete residual analysis for a mixed effects model 
1. Include actual recording time, and check if there is a linear and non-linear effect of time on the observed variables.

```{r loadingdata}
library(arm)
library(lme4)
library(stringi)
library(stringr)
d <- read.csv('one_call_per_row_2020-12-17.csv')
# create a single/multi-bat categorical variable
d['multi_bat'] <- factor(d['num_bats']>1)
d['annot_id'] <- stri_sub(d$video_annot_id,8,-1)
d['annot_id'] <- str_replace_all(d$annot_id, '21502300','2123')
# sort the dataset by time first
time_ordering <- stri_order(d$annot_id,numeric=TRUE)
d_tordered <- d[time_ordering,]

# - is there an influence of the recording hour? 
split.and.join <- function(x){
  splitx <- str_split(x,'_')
  jointx <- paste(splitx[[1]][1],'_',splitx[[1]][2],sep='')
  jointx}
d_tordered$rechour <- sapply(d_tordered$annot_id,split.and.join)

# rescale some parameters to make them more intuitive
d_tordered$tfm_duration <- d_tordered$tfm_duration*10^3 # to ms
d_tordered$ifm_duration <- d_tordered$ifm_duration*10^3 # to ms
d_tordered$cf_duration <- d_tordered$cf_duration*10^3 # to ms

write.csv(d_tordered, 'timeordered_indcall.csv')
```
We will proceed by first running a simple mixed model with random intercept. 
* $parameter$~$group.status + (1|recording.hour)$

### tFM duration ```tFM duration```~```group.status+(1|recording.hour)```

```
The model diagnostics seem mostly okay. Let us also see if there is an effect on recording hour across nights. 

```{r tfmdurn.lmm}
tfmdurn.lmm <- lmer(tfm_duration~multi_bat+(1|rechour), data=d_tordered)
nona.tfm.df <- cbind(d_tordered$multi_bat[!is.na(d$tfm_duration)],resid(tfmdurn.lmm))
colnames(nona.tfm.df) <- c('multi_bat','resid.tfm.durn')
nona.tfm.df <- as.data.frame(nona.tfm.df)
# model diagnostics 
par(mfrow=c(2,2))
boxplot(resid.tfm.durn~multi_bat, data=nona.tfm.df)
qqnorm(resid(tfmdurn.lmm ))
qqline(resid(tfmdurn.lmm ))   
hist(subset(nona.tfm.df, multi_bat==1)$resid.tfm.durn)
hist(subset(nona.tfm.df, multi_bat==2)$resid.tfm.durn)
```


### iFM duration ```iFM duration```~```group.status+(1|recording.hour)```
```{r ifmdurn}
noifmna <- d_tordered[!is.na(d_tordered$ifm_duration),] # remove rows with missing ifm values
ifm.lmm <- lmer(ifm_duration~multi_bat+(1|rechour), data=noifmna)

ifmmodelresid <- as.data.frame(cbind(resid(ifm.lmm),noifmna$multi_bat))
colnames(ifmmodelresid)<-c('resid','multi_bat')
# model diagnostics 
par(mfrow=c(2,2))
boxplot(resid~multi_bat, data=ifmmodelresid)
qqnorm(resid(ifm.lmm ))
qqline(resid(ifm.lmm ))   
hist(subset(ifmmodelresid, multi_bat==1)$resid)
hist(subset(ifmmodelresid, multi_bat==2)$resid)

```
The diagnostics of the iFM model suggest a good fit to the assumptions of the lmm. 

### CF duration ```CF duration```~```group.status+(1|recording.hour)```
```{r cfdurn}
cfdurn.nona <- d_tordered[!is.na(d_tordered$cf_duration),]
cfdurn.lmm <- lmer(cf_duration~multi_bat+(1|rechour), data=cfdurn.nona)

cfmodelresid <- as.data.frame(cbind(resid(cfdurn.lmm),cfdurn.nona$multi_bat))
colnames(cfmodelresid)<-c('resid','multi_bat')
# model diagnostics 
par(mfrow=c(1,3))
#boxplot(resid~multi_bat, data=cfmodelresid)
qqnorm(resid(cfdurn.lmm ))
qqline(resid(cfdurn.lmm))   
hist(subset(cfmodelresid, multi_bat==1)$resid, main='Residuals, single bat')
hist(subset(cfmodelresid, multi_bat==2)$resid, main='Resduals, multi bat')
```
The model fit isn't great. This may hve to do with the dominance of low duration calls, and a few very long calls. What if we transform the cf duration? 

```{r cfdurn.trans}
cfdurn.nona$logcfdurn <- log(cfdurn.nona$cf_duration)
cfdurn2.lmm <- lmer(logcfdurn~multi_bat+(1|rechour), data=cfdurn.nona)

cfmodel2resid <- as.data.frame(cbind(resid(cfdurn2.lmm),cfdurn.nona$multi_bat))
colnames(cfmodel2resid)<-c('resid','multi_bat')
# model diagnostics 
par(mfrow=c(2,2))
#boxplot(resid~multi_bat, data=cfmodelresid)
qqnorm(resid(cfdurn2.lmm ))
qqline(resid(cfdurn2.lmm))   
hist(subset(cfmodel2resid, multi_bat==1)$resid, main='Residuals, single bat')
hist(subset(cfmodel2resid, multi_bat==2)$resid, main='Resduals, multi bat')
plot(ranef(cfdurn2.lmm))
qqline(ranef(cfdurn2.lmm))

```
The fit is much better than before, and the distribution of residuals is also more similar between the two groups. 
